---
title: "Aides Drug"
author: "Ana Mendez"
date: "03/18/2025"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

First we need to take all the data that is in triUmph, and produce a subset of just females. 


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->


```{r}
Females<-filter(triUmp,gender=="F")
```

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected+treatment,data=Females)
```

```{r}
barchartGC(~infected+treatment,data=Females,type = "percent")
```

We have generated two bar charts to compare the data, the first shows total number of females who are involved in the analysis. In chart 1, we see that most females who are drug users do not contract AIDS, however it does not show us much about the effectiveness of the drug. In chart 2, we have a clearer picture because it shows the data as a percentage. In this test we see that the drog is bad.


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1<-xtabs(~infected+treatment,data=triUmp)
rowPerc(table1)
colPerc(table1)
```


 In this version of the results we see the number of people in each group and the rate at which they contracted AIDS in the infected group, we see that there are about the same in the none category, but in the infected category we see that the group who took the drug 59% has more AIDS than the ones who took the placebo 40.7%.


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1) 
fisher.test(table1)
```

We ran three inferential tests in this section, Two Chi-Squared tests, and a Fisher's Exact Test. In the Chi-Squared Test we are getting a p-value of 0.0000002761, the second Chi-squared test gave us 0  which is pretty low, for the Fishers test we got 0.000000233, which again is pretty low. Given this results we reject the null hypotesis in all of the cases

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The odds ration from getting this disease, if you take the drug goes form .63 to .81, meaning that you have more odds of getting the disease if you take this grug, we have an ineffective drug, and we conclude that the drug doesn't work.

As we worked throught this categorical data, we compared two different categories: you got aids and you didnt got adis depending on the results based off of it you took 
